---
title: "Clinical Sequence Analysis"
author: "Stafford"
date: "11/02/2020"
output: html_document
---
#Unknown Sequence Analysis

unknown sequence "Seq" sampled from patient X biofluids was found to be a possible match with the genome of the newly discovered Coronavirus (2019-nCoV), sequence matched with Severe acute respiratory syndrome coronavirus 2

Here is a [link](https://www.who.int/emergencies/diseases/novel-coronavirus-2019/technical-guidance/patient-management) to WHO's technical guide for patient management.

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)

library(BiocManager)
library(genbankr)
library(rentrez)
library(annotate)
library(ggtree)
library(ape)
library(muscle)
library(reshape2)
```

Unknown Sequence - Patient X
```{r sequence_input}
Seq <- c("ATGTCTGATAATGGACCCCAAAATCAGCGAAATGCACCCCGCATTACGTTTGGTGGACCCTCAGATTCAA
CTGGCAGTAACCAGAATGGAGAACGCAGTGGGGCGCGATCAAAACAACGTCGGCCCCAAGGTTTACCCAA
TAATACTGCGTCTTGGTTCACCGCTCTCACTCAACATGGCAAGGAAGACCTTAAATTCCCTCGAGGACAA
GGCGTTCCAATTAACACCAATAGCAGTCCAGATGACCAAATTGGCTACTACCGAAGAGCTACCAGACGAA
TTCGTGGTGGTGACGGTAAAATGAAAGATCTCAGTCCAAGATGGTATTTCTACTACCTAGGAACTGGGCC
AGAAGCTGGACTTCCCTATGGTGCTAACAAAGACGGCATCATATGGGTTGCAACTGAGGGAGCCTTGAAT
ACACCAAAAGATCACATTGGCACCCGCAATCCTGCTAACAATGCTGCAATCGTGCTACAACTTCCTCAAG
GAACAACATTGCCAAAAGGCTTCTACGCAGAAGGGAGCAGAGGCGGCAGTCAAGCCTCTTCTCGTTCCTC
ATCACGTAGTCGCAACAGTTCAAGAAATTCAACTCCAGGCAGCAGTAGGGGAACTTCTCCTGCTAGAATG
GCTGGCAATGGCGGTGATGCTGCTCTTGCTTTGCTGCTGCTTGACAGATTGAACCAGCTTGAGAGCAAAA
TGTCTGGTAAAGGCCAACAACAACAAGGCCAAACTGTCACTAAGAAATCTGCTGCTGAGGCTTCTAAGAA
GCCTCGGCAAAAACGTACTGCCACTAAAGCATACAATGTAACACAAGCTTTCGGCAGACGTGGTCCAGAA
CAAACCCAAGGAAATTTTGGGGACCAGGAACTAATCAGACAAGGAACTGATTACAAACATTGGCCGCAAA
TTGCACAATTTGCCCCCAGCGCTTCAGCGTTCTTCGGAATGTCGCGCATTGGCATGGAAGTCACACCTTC
GGGAACGTGGTTGACCTACACAGGTGCCATCAAATTGGATGACAAAGATCCAAATTTCAAAGATCAAGTC
ATTTTGCTGAATAAGCATATTGACGCATACAAAACATTCCCACCAACAGAGCCTAAAAAGGACAAAAAGA
AGAAGGCTGATGAAACTCAAGCCTTACCGCAGAGACAGAAGAAACAGCAAACTGTGACTCTTCTTCCTGC
TGCAGATTTGGATGATTTCTCCAAACAATTGCAACAATCCATGAGCAGTGCTGACTCAACTCAGGCCTAA")
```


```{r genbank_retrival}
HsBLAST<-blastSequences(paste(Seq),as = 'data.frame',
                           hitListSize = 20, timeout = 400)

HsHitsDF<-data.frame(ID=HsBLAST$Hit_accession,Seq=HsBLAST$Hsp_hseq, stringsAsFactors = FALSE)
```

## Species Hits

The following are the species for which the unknown sequence was found in.

```{r analysis}
HsHitsDNA<-sapply(HsHitsDF$Seq,strsplit,split="")
names(HsHitsDNA)<-paste(1:nrow(HsHitsDF),HsHitsDF$ID,sep="_")

HsHitSeqs<-read.GenBank(HsBLAST$Hit_accession[1:3])
attr(HsHitSeqs, "species")
```

## Alignment of Severe acute respiratory syndrome coronavirus 2 with the unknown sequence query.

```{r sequence_alignment}
HsHitsDNA<-as.DNAbin(HsHitsDNA)
HsAlign<-ape::muscle(HsHitsDNA,quiet=F)

checkAlignment(HsAlign, what=1)
```
Figure 1. Alignment of the unknown sequnce with the sequence documented for Severe acute respiratory syndrome (coronavirus 2)

## Phylogenetic Map from Sequence Alignment

```{r tree, Figure 2. Phylogenetic mapping of Severe acute respiratory syndrome (coronavirus 2) within the coronavirus family}
HsDM<-dist.dna(HsAlign, model="K80")
class(HsDM)

HsTree<-nj(HsDM)
str(HsTree)
class(HsTree)
ggtree(HsTree, branch.length = "none", layout = "circular")+ 
  geom_tiplab(angle=15)
```
